---
title: "Scoring with LLM"
output: 
  html_document:
    theme:
      version: 5
      bootswatch: lux
params:
  analysis_result: !r NULL

---

```{r include=FALSE, error=TRUE}
analysis_result <- params$analysis_result
stopifnot(inherits(analysis_result, "AnalysisResult"))

report_df <- .kwallm_report_results_df(analysis_result)
group_lookup <- .kwallm_report_group_lookup(analysis_result)
result_payload <- analysis_result@results
reliability_summary <- if (is.null(analysis_result@reliability)) {
  NULL
} else {
  analysis_result@reliability@summary
}
main_model <- .kwallm_get_stage_model_id(analysis_result, "scoring")
research_background <- analysis_result@metadata@research_background
by_column_name <- analysis_result@input@grouping_column

source(here::here("R", "style_datatable_config.R"), local = TRUE)
source(here::here("R", "report_grouped_frequencies.R"), local = TRUE)
```

## Method

`r .kwallm_report_text_count_summary(analysis_result)` The LLM (model: `r htmltools::htmlEscape(main_model)`) was asked to provide a score (0-100) indicating
how well a text fits the characteristic '*`r htmltools::htmlEscape(result_payload@characteristic)`*'.

Each text was presented separately to the LLM.

```{r echo=FALSE, results='asis', error=TRUE}
research_background <- htmltools::htmlEscape(research_background)
if (
  !is.null(analysis_result@metadata@research_background)
  && research_background != ""
) {
  cat(paste0(
    "### Research background\n",
    "The user provided the following background information about the research:\n\n",
    "<blockquote><i>",
  research_background,
    "</i></blockquote>", "\n\n",
    "This information was added to each prompt sent to the language model, so that it could better understand the context of the research."
  ))
}
```

```{r echo=FALSE, results='asis', error=TRUE}
if (!is.null(reliability_summary)) {
  sensitivity_sentence <- htmltools::htmlEscape(
    reliability_summary$sensitivity_sentence
  )
  # Guard against NaN values from degenerate t-test (constant/equal scores)
  irr_statistic <- ifelse(
    is.nan(reliability_summary$statistic),
    0,
    reliability_summary$statistic
  )
  irr_p_value <- ifelse(
    is.nan(reliability_summary$p.value),
    1,
    reliability_summary$p.value
  )
  irr_estimate <- ifelse(
    is.nan(reliability_summary$estimate),
    0,
    reliability_summary$estimate
  )
  irr_conf_low <- ifelse(
    is.nan(reliability_summary$conf.low),
    NA,
    reliability_summary$conf.low
  )
  irr_conf_high <- ifelse(
    is.nan(reliability_summary$conf.high),
    NA,
    reliability_summary$conf.high
  )
  cat(paste0(
    "### Reliability\n",
    "The inter-rater reliability between the language model and a human rater was calculated.",
    " A random sample of ",
    reliability_summary$subjects, " texts was taken.",
    "A paired t-test was used to compare the scores given by the language model with those given by the human rater.",
    " The mean difference between the scores is ",
    round(irr_estimate, 2), " (95% CI [",
    round(irr_conf_low, 2), ", ",
    round(irr_conf_high, 2), "];",
    " mean language model = ", round(reliability_summary$llm_mean, 2),
    " [sd = ", round(reliability_summary$llm_sd, 2), "];",
    " mean human rater = ", round(reliability_summary$user_mean, 2),
    " [sd = ", round(reliability_summary$user_sd, 2), "]).",
    " This is a  ",
    dplyr::case_when(
      irr_p_value < 0.05 ~ "significant",
      TRUE ~ "not significant"
    ),
    " difference (t(", reliability_summary$parameter, ") = ",
    round(irr_statistic, 3), ", p = ", round(irr_p_value, 3), ") ",
    " and can be interpreted as ",
    dplyr::case_when(
      irr_p_value >= 0.05 ~ "an agreement between the scores of the language model and the human rater",
      reliability_summary$llm_mean > reliability_summary$user_mean ~ "an overestimation of the scores by the language model",
      reliability_summary$llm_mean < reliability_summary$user_mean ~ "an underestimation of the scores by the language model",
      TRUE ~ "?"
  ), ".\n\n", sensitivity_sentence
  ))
}
```

## Results

### Statistics

Below table shows statistics of the given scores.

```{r results='asis', echo=FALSE, error=TRUE}
report_df |>
  dplyr::summarise(
    Mean = round(mean(result, na.rm = TRUE), 2),
    `Standard deviation` = round(sd(result, na.rm = TRUE), 2),
    Minimum = min(result, na.rm = TRUE),
    Maximum = max(result, na.rm = TRUE),
    Median = median(result, na.rm = TRUE)
  ) |>
  DT::datatable(
    rownames = FALSE,
    extensions = 'Buttons',
    options = get_datatable_options_en()
  )
```

```{r echo=FALSE, results='asis', error=TRUE}
# Show grouped statistics if by_column is specified
if (!is.null(by_column_name) && !is.null(group_lookup)) {
  cat("\n### Statistics per group\n\n")
  cat(paste0(
    "Below table shows score statistics, ",
    "broken down by **", htmltools::htmlEscape(by_column_name), "**.\n\n"
  ))
}
```

```{r echo=FALSE, results='asis', error=TRUE}
if (!is.null(by_column_name) && !is.null(group_lookup)) {
  generate_grouped_score_table(
    df = report_df,
    by_values = group_lookup,
    by_column_name = by_column_name,
    language = "en"
  )
}
```

### Per text

Below table shows the scores per text.

```{r results='asis', echo=FALSE, error=TRUE}
report_df |>
  dplyr::select(result, text) |>
  dplyr::rename(Text = text, Score = result) |>
  dplyr::arrange(dplyr::desc(Score)) |>
  DT::datatable(
    extensions = 'Buttons',
    options = get_datatable_options_en()
  )
```
 
```{r footer, echo=FALSE, results='asis', error=TRUE}
htmltools::tags$footer(
  style = "text-align:center; font-size: 0.8em; color: #888; margin-top: 50px; padding-top: 20px; border-top: 1px solid #eee;",
  htmltools::HTML(paste0(
    "This analysis was performed by KWALLM, an open-source app for automated qualitative text analysis.<br>Developed by ",
    "<a href='https://www.kennispunttwente.nl' target='_blank'>Kennispunt Twente</a> in collaboration with GGD Twente.<br>",
    "Visit the <a href='https://github.com/kennispunttwente/KWALLM' target='_blank'>GitHub-repository</a> for more information.",
    "<br>",
    "<br>",
    format(Sys.Date(), "%Y-%m-%d")
  ))
)
```
