---
title: "Marking with LLM"
output: 
  html_document:
    theme:
      version: 5
      bootswatch: lux
params:
  analysis_result: !r NULL
---

```{r include=FALSE, error=TRUE}
analysis_result <- params$analysis_result
stopifnot(inherits(analysis_result, "AnalysisResult"))

report_df <- .kwallm_report_results_df(analysis_result)
group_lookup <- .kwallm_report_group_lookup(analysis_result)
result_payload <- analysis_result@results
codes <- result_payload@codes$code_text
paragraphs_df <- analysis_result@paragraphs@paragraphs
has_paragraphs <- nrow(paragraphs_df) > 0
paragraph_subject_lookup <- .kwallm_paragraph_subject_lookup(
  analysis_result
)
main_model <- .kwallm_get_stage_model_id(analysis_result, "marking")
research_background <- analysis_result@metadata@research_background
by_column_name <- analysis_result@input@grouping_column

source(here::here("R", "style_datatable_config.R"), local = TRUE)
source(here::here("R", "utils_quotes.R"), local = TRUE)
source(here::here("R", "report_grouped_frequencies.R"), local = TRUE)
```

## Method

`r .kwallm_report_text_count_summary(analysis_result)` The LLM (model: `r htmltools::htmlEscape(main_model)`) was asked to mark text spans that belonged to one of the following codes:

```{r echo=FALSE, results='asis', error=TRUE}
cat("<blockquote>")
cat(paste0("* *", htmltools::htmlEscape(codes), "*"), sep = "\n")
cat("</blockquote>")
```

```{r echo=FALSE, results='asis', error=TRUE}
if (has_paragraphs) {
  cat(paste0(
    "With the texts that were marked, a language model (",
    htmltools::htmlEscape(main_model), ") has finally written a report.", "\n",
    "For each code, a summary of what was said is provided in the report."
  ))
}
```

```{r echo=FALSE, results='asis', error=TRUE}
research_background <- htmltools::htmlEscape(research_background)
if (
  !is.null(analysis_result@metadata@research_background)
  && research_background != ""
) {
  cat(paste0(
    "### Research background\n",
    "The user provided the following background information about the research:\n\n",
    "<blockquote><i>",
  research_background,
    "</i></blockquote>", "\n\n",
    "This information was added to each prompt sent to the language model, so that it could better understand the context of the research."
  ))
}
```

## Results

```{r echo=FALSE, results='asis', error=TRUE}
if (!is.null(by_column_name) && !is.null(group_lookup)) {
  cat("### Code prevalence by group\n\n")
  cat(paste0(
    "Below table shows, for each code, the percentage of texts within each **",
    htmltools::htmlEscape(by_column_name),
    "** group where that code was marked at least once.\n\n"
  ))
}
```

```{r echo=FALSE, results='asis', error=TRUE}
if (!is.null(by_column_name) && !is.null(group_lookup)) {
  generate_grouped_marking_prevalence_table(
    df = report_df,
    by_values = group_lookup,
    by_column_name = by_column_name,
    codes = codes,
    language = "en"
  )
}
```

```{r echo=FALSE, results='asis', error=TRUE}
if (!is.null(by_column_name) && !is.null(group_lookup)) {
  cat("\n### Texts with marked spans per group\n\n")
  cat(paste0(
    "Below table shows, per group and code, how many texts contained at least one marked span, ",
    "plus the total number of marked spans that were returned.\n\n"
  ))
}
```

```{r echo=FALSE, results='asis', error=TRUE}
if (!is.null(by_column_name) && !is.null(group_lookup)) {
  generate_grouped_marking_frequency_table(
    df = report_df,
    by_values = group_lookup,
    by_column_name = by_column_name,
    codes = codes,
    language = "en"
  )
}
```

### Marked texts

Below table shows, for each text row, chunk, and code, which part of the text was marked by the language model.

Column 'Text' contains the analysis text (after any preprocessing or anonymization), column 'Chunk' contains the (possibly preprocessed) chunk that was shown to the language model,
the column 'Code' contains the code for which the text was checked, the column 'Marked text' contains the matched substring from the source text (which may differ slightly from the model's raw output when fuzzy recovery was used to snap the match to the actual text),
and the column 'Response status' indicates whether the chunk/code response fully resolved or only partial literal matches survived the correction loop.
An empty value in 'Marked text' means no relevant span was returned for that chunk/code combination.

```{r, echo=FALSE, results='asis', error=TRUE}

report_df |>
  dplyr::mutate(marked_text = dplyr::coalesce(marked_text, "")) |>
  dplyr::rename(
    `Text` = text,
    `Chunk` = chunk_text,
    `Code` = code,
    `Marked text` = marked_text,
    `Response status` = response_status
  ) |>
  DT::datatable(
    rownames = FALSE,
    extensions = 'Buttons',
    options = get_datatable_options_en()
  )

```

```{r echo=FALSE, results='asis', error=TRUE}
if (!has_paragraphs) return(invisible(NULL))

output <- list()

output <- append(output, list(
  htmltools::HTML("<h3>Report</h3>"),
  htmltools::HTML("<p>The following paragraphs are summaries of the marked texts per code.</p>"),
  htmltools::div(
    class = "d-flex justify-content-center",
    htmltools::div(
      class = "d-flex justify-content-center",
      htmltools::div(
        class = "alert alert-warning p-2 text-center",
        style = "max-width: 500px; font-size: 0.9rem;",
        htmltools::tags$div(
          class = "d-flex align-items-center",
          bsicons::bs_icon("exclamation-triangle-fill", class = "me-2 fs-1"),
          htmltools::tags$small(
            htmltools::tags$i(
              "Language models can 'hallucinate'.",
              " Check whether the summaries are supported by the texts.",
              " Icons next to quotes indicate whether they appear in the texts."
            )
          )
        )
      )
    )
  )
))

## quote utils loaded from report_quote_utils.R

for (i in seq_len(nrow(paragraphs_df))) {
  paragraph_row <- paragraphs_df[i, , drop = FALSE]
  topic <- unname(paragraph_subject_lookup[[
    as.character(paragraph_row$subject_id[[1]])
  ]])
  if (is.null(topic) || !length(topic)) {
    topic <- ""
  }
  topic <- htmltools::htmlEscape(topic)
  paragraph_text <- htmltools::htmlEscape(paragraph_row$paragraph_text[[1]])
  
  supporting_texts <- .kwallm_paragraph_supporting_texts(
    analysis_result,
    paragraph_row$paragraph_id[[1]]
  )
  if (!length(supporting_texts)) {
    next
  }
  #### 1 Verify quotes on presence in supporting texts ####
  paragraph_text <- verify_and_decorate_quotes(
    paragraph_text = paragraph_text,
    supporting_texts = supporting_texts,
    lang = "en",
    escape_html = FALSE
  )
  
  #### 2 Check if paragraph has attribute about prompt not fitting in context window ####
  
  # If paragraph has attribute 'prompt_fits' and it is FALSE, add a warning
  #   to the output
  warning_context_window <- NULL
  prompt_fits <- isTRUE(paragraph_row$prompt_fits[[1]])
  if (isFALSE(prompt_fits)) {
    warning_context_window <- htmltools::div(
      class = "d-flex justify-content-center",
      htmltools::div(
        class = "d-flex justify-content-center",
        htmltools::div(
          class = "alert alert-warning p-2 text-center",
          style = "max-width: 500px; font-size: 0.9rem;",
          htmltools::tags$div(
            class = "d-flex align-items-center",
            bsicons::bs_icon("exclamation-triangle-fill", class = "me-2 fs-1"),
            htmltools::tags$small(
              htmltools::tags$i(
                "The prompt for the summary of this code did",
                " not fit in the context window of the language model.",
                " It is possible that not all instructions or texts were seen by the language model."
              )
            )
          )
        )
      )
    )
  }
  
  # Append processed output
  output <- append(output, list(
  htmltools::HTML(paste0("<h4>", topic, "</h4>")),
    warning_context_window,
  htmltools::HTML(paste0("<p>", paragraph_text, "</p>"))
  ))

  if (length(supporting_texts)) {
    # Convert **bold** syntax to <strong>bold</strong>
    formatted_texts <- lapply(supporting_texts, function(txt) {
      # Replace **text** with <strong>text</strong>
      htmltools::HTML(gsub(
        "\\*\\*(.*?)\\*\\*",
        "<span style='font-weight: 900; color: #b30000; background-color: #ffcccc; text-decoration: underline;'>\\1</span>",
        htmltools::htmlEscape(txt)
      ))
    })
  
    # Create list items
    list_items <- lapply(formatted_texts, function(txt) {
      paste0("<li><i>", txt, "</i></li>")
    })
  
    acc <- bslib::accordion(
      bslib::accordion_panel(
        title = "View texts",
        htmltools::HTML(paste0("<ul>", paste0(list_items, collapse = ""), "</ul>"))
      ),
      id = paste0("acc-", i),
      open = FALSE
    )
    
    output <- append(output, list(acc))
  }

  output <- append(output, list(htmltools::HTML("<br>")))
}

if (nrow(paragraphs_df) > 0) {
  htmltools::tagList(output)
}
``` 
 
```{r footer, echo=FALSE, results='asis', error=TRUE}
htmltools::tags$footer(
  style = "text-align:center; font-size: 0.8em; color: #888; margin-top: 50px; padding-top: 20px; border-top: 1px solid #eee;",
  htmltools::HTML(paste0(
    "This analysis was performed by KWALLM, an open-source app for automated qualitative text analysis.<br>Developed by ",
    "<a href='https://www.kennispunttwente.nl' target='_blank'>Kennispunt Twente</a> in collaboration with GGD Twente.<br>",
    "Visit the <a href='https://github.com/kennispunttwente/KWALLM' target='_blank'>GitHub-repository</a> for more information.",
    "<br>",
    "<br>",
    format(Sys.Date(), "%Y-%m-%d")
  ))
)
```
