---
title: "Topic extraction with LLM"
output: 
  html_document:
    theme:
      version: 5
      bootswatch: lux
params:
  analysis_result: !r NULL
---

```{r include=FALSE, error=TRUE}
library(stringr)
analysis_result <- params$analysis_result
stopifnot(inherits(analysis_result, "AnalysisResult"))

report_df <- .kwallm_report_results_df(analysis_result)
group_lookup <- .kwallm_report_group_lookup(analysis_result)
result_payload <- analysis_result@results
assign_multiple_categories <- isTRUE(result_payload@multi_label)
reliability_summary <- if (is.null(analysis_result@reliability)) {
  NULL
} else {
  analysis_result@reliability@summary
}
paragraphs_df <- .kwallm_report_paragraphs_by_frequency(analysis_result)
has_paragraphs <- nrow(paragraphs_df) > 0
paragraph_subject_lookup <- .kwallm_paragraph_subject_lookup(
  analysis_result
)
main_model <- .kwallm_get_stage_model_id(
  analysis_result,
  c("topic_candidate_generation", "topic_assignment")
)
reduction_model <- .kwallm_get_stage_model_id(
  analysis_result,
  c("topic_reduction", "topic_not_applicable_check")
)
research_background <- analysis_result@metadata@research_background
by_column_name <- analysis_result@input@grouping_column
topic_provenance <- result_payload@topic_provenance
topic_batch_size <- topic_provenance@batch_size
topic_draws <- topic_provenance@draws
topic_n_batches <- topic_provenance@n_batches
topics_human_edited <- isTRUE(topic_provenance@human_edited)
auto_added_not_applicable <- isTRUE(topic_provenance@auto_added_not_applicable)

source(here::here("R", "style_datatable_config.R"), local = TRUE)
source(here::here("R", "utils_quotes.R"), local = TRUE)
source(here::here("R", "report_grouped_frequencies.R"), local = TRUE)
```

```{r echo=FALSE, results='asis', error=TRUE}
analysis_name <- analysis_result@metadata@analysis_name
if (nzchar(analysis_name)) {
  cat(paste0(
    "<h3 style='color: #666; margin-top: -0.5em;'>",
    htmltools::htmlEscape(analysis_name),
    "</h3>\n"
  ))
}
```

## Method

`r .kwallm_report_text_count_summary(analysis_result)` A language model was then used
to discover which different topics are mentioned in the texts
and how often.

A topic modeling procedure has been applied based on
[Wanrooij, Manhar, and Yang (2024)](https://bnaic2024.sites.uu.nl/wp-content/uploads/sites/986/2024/10/Topic-Modeling-for-Small-Data-using-Generative-LLMs.pdf)
and [Pham et al. (2023)](https://arxiv.org/abs/2311.01449). This procedure consists of 3 steps:

```{r echo=FALSE, results='asis', error=TRUE}
step_1_text <- paste0(
  "* In step 1, the unique analysis-unit texts are grouped into prompt batches"
)
if (!is.null(topic_batch_size)) {
  step_1_text <- paste0(
    step_1_text,
    " of up to ",
    topic_batch_size,
    " texts"
  )
}
step_1_text <- paste0(
  step_1_text,
  ". The first draw preserves the upload order; any additional draws reshuffle the texts."
)
if (!is.null(topic_draws)) {
  step_1_text <- paste0(
    step_1_text,
    " In this run, each text could be drawn up to ",
    topic_draws,
    if (topic_draws == 1) " time." else " times."
  )
}
if (!is.null(topic_n_batches)) {
  step_1_text <- paste0(
    step_1_text,
    " This produced ",
    topic_n_batches,
    " candidate-topic batches."
  )
}
step_1_text <- paste0(
  step_1_text,
  " The model was asked to respond with possible topics that are neither too specific nor too general. This forms the list of candidate topics. For this step, the model ",
  htmltools::htmlEscape(main_model),
  " was used."
)
cat(step_1_text)
```

* In step 2, the list of possible topics
is presented to the LLM, asking it to reduce the list of topics
to a final list without duplications and also avoiding overly specific
or overly general topics. This forms the list of final topics.
For this step, the model `r htmltools::htmlEscape(reduction_model)` has been used.

```{r echo=FALSE, results='asis', error=TRUE}  
multiple_categories_text <- "The LLM was required to assign one topic per text"
if (assign_multiple_categories) {
  multiple_categories_text <- "The LLM was allowed to assign multiple topics per text (at least one topic per text)"
}
```

* In step 3, each text is presented individually to the LLM,
along with the list of final topics, asking it to assign these topics to the text.
`r multiple_categories_text`. For this step, the model `r htmltools::htmlEscape(main_model)` has been used.

```{r echo=FALSE, results='asis', error=TRUE}
if (topics_human_edited || auto_added_not_applicable) {
  cat("### Topic list provenance\n")
  if (topics_human_edited) {
    cat(
      "The final topic list was edited manually before topic assignment.\n\n"
    )
  }
  if (auto_added_not_applicable) {
    cat(
      "A generic 'Unknown/not applicable' topic was added automatically because the reduced topic list did not already contain an equivalent topic.\n\n"
    )
  }
}
```

```{r echo=FALSE, results='asis', error=TRUE}
if (has_paragraphs) {
  cat(paste0(
    "With the texts to which topics were assigned, a language model ",
  "(", htmltools::htmlEscape(main_model), ") finally wrote a report.", "\n",
    "For each topic, a summary was created with quotes summarizing what was said."
  ))
}
```

```{r echo=FALSE, results='asis', error=TRUE}
research_background <- htmltools::htmlEscape(research_background)
if (
  !is.null(analysis_result@metadata@research_background)
  && research_background != ""
) {
  cat(paste0(
    "### Research background\n",
    "The user provided the following background information about the research:\n\n",
    "<blockquote><i>",
  research_background,
    "</i></blockquote>", "\n\n",
    "This information was added to each prompt sent to the language model, so that it could better understand the context of the research."
  ))
}
```

```{r echo=FALSE, results='asis', error=TRUE}
if (!is.null(reliability_summary)) {
  cat(paste0(
    "### Reliability\n",
    "The inter-rater reliability between the language model and a human rater was calculated. ",
    "A random sample of ",
    reliability_summary$subjects,
    if (assign_multiple_categories) " text-topic decisions was taken." else " texts was taken.",
    " The calculated inter-rater reliability is ",
    round(reliability_summary$value, 3), " (", reliability_summary$irr.name, "; ", reliability_summary$stat.name, " = ",
    round(reliability_summary$statistic, 3), ", p = ", round(reliability_summary$p.value, 3), ").",
    " This can be interpreted as a ",
    dplyr::case_when(
      reliability_summary$value < 0 ~ "poor",
      reliability_summary$value < 0.2 ~ "slight",
      reliability_summary$value < 0.4 ~ "fair",
      reliability_summary$value < 0.6 ~ "moderate",
      reliability_summary$value < 0.8 ~ "substantial",
      reliability_summary$value < 1 ~ "almost perfect",
      TRUE ~ "perfect"
    ),
    " agreement between the language model and the human rater ([Landis & Koch, 1977](https://doi.org/10.2307/2529310))."
  ))
}
```

## Results

### Frequency

Below table shows the topics that have been identified and how often
they were mentioned in the texts.

```{r, echo=FALSE, results='asis', error=TRUE}

if (!assign_multiple_categories) {
  report_df |>
  dplyr::group_by(result) |>
  dplyr::summarise(
    Number = dplyr::n(),
    Percentage = round(dplyr::n() / nrow(report_df) * 100, 2)
  ) |>
  dplyr::arrange(dplyr::desc(Number)) |>
  dplyr::rename(Topic = result) |>
  DT::datatable(
    rownames = FALSE,
    extensions = 'Buttons',
    options = get_datatable_options_en(additional_options = list(fixedHeader = TRUE))
  )
} else {
  report_df |>
    dplyr::select(-dplyr::any_of(c("document_id", "text"))) |>
    # For each binary category column, count the number of TRUE values
    dplyr::summarise(dplyr::across(dplyr::everything(), ~ sum(.x, na.rm = TRUE))) |>
    tidyr::pivot_longer(
      cols = dplyr::everything(),
      names_to = "Topic",
      values_to = "Number"
    ) |>
    dplyr::mutate(
      Percentage = round(Number / nrow(report_df) * 100, 2)
    ) |>
    dplyr::arrange(dplyr::desc(Number)) |>
    DT::datatable(
      rownames = FALSE,
      extensions = 'Buttons',
      options = get_datatable_options_en()
    )
}


```

```{r echo=FALSE, results='asis', error=TRUE}
# Show grouped frequencies if by_column is specified
if (!is.null(by_column_name) && !is.null(group_lookup)) {
  cat("\n### Topic prevalence by group\n\n")
  cat(paste0(
    "Below table shows, for each topic, the percentage of texts within each **",
    htmltools::htmlEscape(by_column_name),
    "** group to which that topic was assigned."
  ))

  if (assign_multiple_categories) {
    cat(
      " Because multiple topics can be assigned to one text, percentages across topics can sum to more than 100% within a group.\n\n"
    )
  } else {
    cat("\n\n")
  }
}
```

```{r echo=FALSE, results='asis', error=TRUE}
if (!is.null(by_column_name) && !is.null(group_lookup)) {
  if (!assign_multiple_categories) {
    generate_grouped_topic_prevalence_table_single(
      df = report_df,
      by_values = group_lookup,
      by_column_name = by_column_name,
      topics = result_payload@labels$label_text,
      language = "en"
    )
  } else {
    topic_cols <- setdiff(names(report_df), c("document_id", "text"))
    generate_grouped_topic_prevalence_table_multi(
      df = report_df,
      by_values = group_lookup,
      by_column_name = by_column_name,
      topics = topic_cols,
      language = "en"
    )
  }
}
```

```{r echo=FALSE, results='asis', error=TRUE}
if (!is.null(by_column_name) && !is.null(group_lookup)) {
  cat("\n### Frequency per group\n\n")
  cat(paste0(
    "Below table shows the raw topic counts and within-group percentages, ",
    "broken down by **", htmltools::htmlEscape(by_column_name), "**.\n\n"
  ))
}
```

```{r echo=FALSE, results='asis', error=TRUE}
if (!is.null(by_column_name) && !is.null(group_lookup)) {
  if (!assign_multiple_categories) {
    generate_grouped_topic_table_single(
      df = report_df,
      by_values = group_lookup,
      by_column_name = by_column_name,
      categories = result_payload@labels$label_text,
      language = "en"
    )
  } else {
    # Get topic columns (excluding report identity columns)
    topic_cols <- setdiff(names(report_df), c("document_id", "text"))
    generate_grouped_topic_table_multi(
      df = report_df,
      by_values = group_lookup,
      by_column_name = by_column_name,
      categories = topic_cols,
      language = "en"
    )
  }
}
```

### Per text

Below table shows the assigned topic for each text.

```{r, echo=FALSE, results='asis', error=TRUE}
if (!assign_multiple_categories) {
  report_df |>
  dplyr::select(result, text) |>
  dplyr::rename(Text = text, Topic = result) |>
  dplyr::arrange(dplyr::desc(Topic)) |>
  DT::datatable(
    extensions = 'Buttons',
    options = get_datatable_options_en(additional_options = list(fixedHeader = TRUE))
  )
} else {
  report_df |>
    dplyr::rename(Text = text) |>
    dplyr::arrange(dplyr::desc(Text)) |>
    DT::datatable(
      options = get_datatable_options_en()
    )
}
```
 
```{r echo=FALSE, results='asis', error=TRUE}
if (!has_paragraphs) return(invisible(NULL))

output <- list()

output <- append(output, list(
  htmltools::HTML("<h3>Report</h3>"),
  htmltools::HTML("<p>The following paragraphs are the summaries of the texts per topic.</p>"),
  htmltools::div(
    class = "d-flex justify-content-center",
    htmltools::div(
      class = "d-flex justify-content-center",
      htmltools::div(
        class = "alert alert-warning p-2 text-center",
        style = "max-width: 500px; font-size: 0.9rem;",
        htmltools::tags$div(
          class = "d-flex align-items-center",
          bsicons::bs_icon("exclamation-triangle-fill", class = "me-2 fs-1"),
          htmltools::tags$small(
            htmltools::tags$i(
              "Language models can 'hallucinate'.",
              " Check whether the summaries are supported by the texts.",
              " Icons next to quotes indicate whether they appear in the texts."
            )
          )
        )
      )
    )
  )
))

## quote utils loaded from report_quote_utils.R

for (i in seq_len(nrow(paragraphs_df))) {
  paragraph_row <- paragraphs_df[i, , drop = FALSE]
  topic <- unname(paragraph_subject_lookup[[
    as.character(paragraph_row$subject_id[[1]])
  ]])
  if (is.null(topic) || !length(topic)) {
    topic <- ""
  }
  topic <- htmltools::htmlEscape(topic)
  paragraph_text <- htmltools::htmlEscape(paragraph_row$paragraph_text[[1]])
  
  supporting_texts <- .kwallm_paragraph_supporting_texts(
    analysis_result,
    paragraph_row$paragraph_id[[1]]
  )
  if (!length(supporting_texts) && isTRUE(paragraph_row$prompt_fits[[1]])) {
    next
  }
  #### 1 Verify quotes on presence in supporting texts ####
  if (length(supporting_texts)) {
    paragraph_text <- verify_and_decorate_quotes(
      paragraph_text = paragraph_text,
      supporting_texts = supporting_texts,
      lang = "en",
      escape_html = FALSE
    )
  }
  
  #### 2 Check if paragraph has attribute about prompt not fitting in context window ####
  
  # If paragraph has attribute 'prompt_fits' and it is FALSE, add a warning
  #   to the output
  warning_context_window <- NULL
  prompt_fits <- isTRUE(paragraph_row$prompt_fits[[1]])
  if (isFALSE(prompt_fits)) {
    warning_context_window <- htmltools::div(
      class = "d-flex justify-content-center",
      htmltools::div(
        class = "d-flex justify-content-center",
        htmltools::div(
          class = "alert alert-warning p-2 text-center",
          style = "max-width: 500px; font-size: 0.9rem;",
          htmltools::tags$div(
            class = "d-flex align-items-center",
            bsicons::bs_icon("exclamation-triangle-fill", class = "me-2 fs-1"),
            htmltools::tags$small(
              htmltools::tags$i(
                "The prompt for the summary of this topic did",
                " not fit in the context window of the language model.",
                " It is possible that not all instructions or texts were seen by the language model."
              )
            )
          )
        )
      )
    )
  }

  warning_sampled_sources <- NULL
  if (identical(paragraph_row$source_coverage[[1]], "sampled")) {
    warning_sampled_sources <- htmltools::div(
      class = "alert alert-warning p-2 text-center",
      style = "max-width: 500px; margin-left: auto; margin-right: auto; font-size: 0.9rem;",
      htmltools::tags$small(htmltools::tags$i(
        "Because all texts did not fit in the model context window, this",
        " summary is based on a random context-sized sample of the topic texts."
      ))
    )
  }
  
  # Append processed output
  output <- append(output, list(
  htmltools::HTML(paste0("<h4>", topic, "</h4>")),
    warning_context_window,
    warning_sampled_sources,
    htmltools::HTML(paste0("<p>", paragraph_text, "</p>"))
  ))

  if (length(supporting_texts)) {
  acc <- bslib::accordion(
      bslib::accordion_panel(
        title = "View texts",
        htmltools::HTML(paste0("<ul>", paste0("<li><i>", htmltools::htmlEscape(supporting_texts), "</i></li>", collapse = ""), "</ul>")),
      ),
      id = paste0("acc-", i),
      open = FALSE
    )
    output <- append(output, list(acc))
  }

  output <- append(output, list(htmltools::HTML("<br>")))
}

if (nrow(paragraphs_df) > 0) {
  htmltools::tagList(output)
}
``` 

```{r footer, echo=FALSE, results='asis', error=TRUE}
htmltools::tags$footer(
  style = "text-align:center; font-size: 0.8em; color: #888; margin-top: 50px; padding-top: 20px; border-top: 1px solid #eee;",
  htmltools::HTML(paste0(
    "This analysis was performed by KWALLM, an open-source app for automated qualitative text analysis.<br>Developed by ",
    "<a href='https://www.kennispunttwente.nl' target='_blank'>Kennispunt Twente</a> in collaboration with GGD Twente.<br>",
    "Visit the <a href='https://github.com/kennispunttwente/KWALLM' target='_blank'>GitHub-repository</a> for more information.",
    "<br>",
    "<br>",
    format(Sys.Date(), "%Y-%m-%d")
  ))
)
```
